\cardfrontfoot{Week 5: \textit{Drosophila melanogaster II, gap genes}}

\begin{flashcard}[Gap genes, anterior-posterior]{The three important anterior-posterior gap genes are ..., and they are repressed by ...}
    Krüppel, repressed by all except bicoid (knirps, giant, hunchback, tailless, etc.) \\
    knirps, repressed by all except bicoid \\
    giant, repressed by all except bicoid \textbf{and hunchback}
    \textbf{All transcription factors.}
\end{flashcard}

\begin{flashcard}[Gap genes, anterior-posterior]{If an anterior-posterior gap gene is deleted ...}
    The region normally specified is deleted.
\end{flashcard}

\begin{flashcard}[Gap genes, anterior-posterior]{Gap genes ... of type ... are required for acron and activated by ...}
    Bicoid actiavtes otd, ems and btd, \textbf{only} required for acron, while tailless and huckebein from torso are required for acron and telson. They are all \textbf{transcription factors}
\end{flashcard}

\begin{flashcard}[Gap genes, dorsal-ventral]{The dorsal side is initially specified by ...}
    After anterior-posterior specification, nucleus releases Gurken mRNA at random side - \textbf{dorsal side}. Gurken activates torpedo, which inhibits pipe, which is used in the ventral side. 
\end{flashcard}

\begin{flashcard}[Gap genes, dorsal-ventral]{The ventral side reaction chain involves the proteins..}
    Nudel and pipe activate proteases Gd --> Snake --> Easter. Easter activates ligand Spätzle, which binds to membrane receptor Toll. Toll releases Cactus from Dorsal, allowing Dorsal into the nucleus, specifying ventral side. 
\end{flashcard}


\begin{flashcard}[Gap genes, dorsal-ventral]{A fly is ventralised in the case of a LOF in proteins..}
    Cactus, as it inhibits dorsal. Gurken and torpedo, as they inhibit pipe
\end{flashcard}

\begin{flashcard}[Gap genes, dorsal-ventral]{A fly is dorsalised in the case of a LOF in proteins..}
    Pipe, Nudel, Gd, Snake, Easter, Spätzle, Toll and Dorsal.
\end{flashcard}

\begin{flashcard}[Dorsal concentration gradient]{Highest concentrations of dorsal activate ... which specify the ...}
    Proteins twist and snail, specifies the mesoderm (e.g. gut)
\end{flashcard}

\begin{flashcard}[Dorsal concentration gradient]{Low concentrations of dorsal activate ... (top to bottom)}
    rhomboid, short gastrulation and single-minded (lowest affinity for dorsal of the three) 
\end{flashcard}

\begin{flashcard}[Dorsal concentration gradient]{Dorsal represses ... (top to bottom)}
    zerknüllt, tolloid, decapentaplegic.
\end{flashcard}

\begin{flashcard}[Dorsal concentration gradient]{Snail and twist are not activated at low dorsal concentrations due to ...}
    Low activator site affinity for dorsal. Harder to bind long enough for transcription to start. 
\end{flashcard}




\begin{flashcard}[Segment polarity genes]{Segment polarity genes are activated by ... and the involved genes are ...}
    Engrailed activated by even skipped and fushi taratzu. Involved genes:
    Engrailed --> hedgehog --| patched --| smoothened --> Cubitus Interruptus --> wingless --> frizzled --| zeste white 3 --| armadillo --> engrailed
    Engrailed and cubitus interruptus are TFs. 
\end{flashcard}

\begin{flashcard}[Pair rule genes]{The pair rule genes are initially in the ... and move to the ... in the segments}
    anterior --> posterior.
\end{flashcard}

\begin{flashcard}[RNAi vs genetic screening]{The benefits of RNAi over genetic screening and vice versa are...}
    \textbf{Genetic screening}: Can discover tissue-specific promotorers by non-coding DNA mutations. Can discover gain of function mutations, though lethal genes hard to mutate
    \textbf{RNAi} Much cheaper, bypasses lethal gene limits, discovers redundant genes.
\end{flashcard}

\begin{flashcard}[Pair rule genes]{Even-skipped stripe two is activated by ... and repressed by ... }
    Activated by hunchback and bicoid. Repressed to the anterior by giant, to the posterior by Krüppel.
\end{flashcard}

\begin{flashcard}[Homeotic genes]{Homeotic genes and segment polarity genes appear after ...}
    The cellular blastoderm (individual cells) has formed.
\end{flashcard}

\begin{flashcard}[Homeotic genes]{Homeotic genes do ...}
    Homeotic genes are localized genes that activate realizator genes, leading to the local development of appendages, etc. 
\end{flashcard}

\begin{flashcard}[Homeotic genes]{Antennapedia LOF causes... \\ GOF causes...}
    LOF causes antennae to appear at 2nd legs, along with other thorax modifications \\
    GOF causes antennae to be replaced by legs.
\end{flashcard}

\begin{flashcard}[Homeotic genes]{Homeotic genes resemble the ... genes in humans}
    HOX-genes.
\end{flashcard}

\begin{flashcard}[Homeotic genes]{In homeotic genes, LOF causes ... while GOF causes ...}
    LOF causes anteriorizations, GOF causes posteriorizations due to posterior dominance. Posterization = anterior part gets posterior elements, e.g. legs instead of antennae.
\end{flashcard}

\begin{flashcard}[Homeotic genes]{Antennapedia GOF replaces antennae with legs due to ...}
    Antennapedia represses the antennae-related genes and activates leg-related genes.
\end{flashcard}

\begin{flashcard}[Homeotic genes]{The three last homeotic genes are ... and they are each necessary for ..}
    Ultrabitthorax, abdominal A and abdominal B. Ubx controls T3-A1, abdA A2-A4 and abdB A5-A8. Each required for their part, posterior genes take precedence.
\end{flashcard}
